DMRGenie: Graphical interface for entanglement renormalization

### 2025-12-03 paper/paper.tex
Add files via upload

Methods of entanglement renormalization use quantities native to quantum information ({\it i.e.}, the density matrix or the entropy) in order to optimize a a quantum problem to find solutions to the eigenvalue problem \cite{townsend2000modern}. Entanglement renormalization techniques are formulated on a graph of tensors and have definite rules for tensor operations. Because tensor computations and operations can be cumbersome to represent mathematically in index notation, the field has adopted a diagram notation that allows for the improved communication of algorithms. Namely, the rank of the tensor corresponds to the number of lines on the diagram \cite{penrose1971angular}. An example is shown for rank-1, 2, and 3 tensors in Fig.~\ref{diagram}.

@@

The use of the diagram notation is useful for representing large numbers of tensors in a coherent way without representing the full mathematical detail (see Fig.~\ref{network}). For example, the wavefunction when decomposed through a series of reshapes and singular value decompositions (SVD) takes the form \cite{bakerCJP21}
\begin{align}\nonumber
|\psi\rangle&=\sum_{\substack{\sigma_1\sigma_2\sigma_3\\\sigma_4\sigma_5\sigma_6}}\sum_{\substack{a_1a_2a_3\\a_4a_5}}A^{\sigma_1}_{a_1}A^{\sigma_2}_{a_1a_2}A^{\sigma_3}_{a_2a_3}A^{\sigma_4}_{a_3a_4}A^{\sigma_5}_{a_4a_5}A^{\sigma_6}_{a_5}\\
&\quad\quad\quad\quad\quad\quad\quad\quad\times|\sigma_1\sigma_2\sigma_3\sigma_4\sigma_5\sigma_6\rangle
\label{mpsform}
\end{align}
which is far more intricate than the diagram notation but conveys no extra information. Any quantum operator (for example, the Hamiltonian) can be decomposed into a similar site-by-site form involving many tensors.

@@

The library implements several classes of tensor network anstaz to represent wavefunctions of quantum problems. This includes the matrix product state (MPS) \cite{verstraete2006matrix,baker2024bundled}, projected entangled pair states (PEPS) \cite{verstraete2004renormalization}, multiscale entanglement renormalization anstaz (MERA) \cite{vidal2007entanglement}, and classical tensor network algorithms such as the tensor renormalization group (TRG) \cite{levin2007tensor}. Each of these implementations is done with care towards ease of use, culminating in the DMRGenie implementation here.  All of these tensor networks are implemented in a fully flexible way.

>>>>

Methods of entanglement renormalization use quantities native to quantum information ({\it i.e.}, the density matrix or the entropy) in order to optimize a a quantum problem to find solutions to the eigenvalue problem \cite{townsend2000modern}. Entanglement renormalization techniques are formulated on a graph of tensors and have definite rules for tensor operations. Because tensor computations and operations can be cumbersome to represent mathematically in index notation, the field has adopted a diagram notation that allows for the improved communication of algorithms. Namely, the rank of the tensor corresponds to the number of lines attached to a shape on the diagram \cite{penrose1971angular}. An example is shown for rank-1, 2, and 3 tensors in Fig.~\ref{diagram}.

@@

The use of the diagram notation is useful for representing large numbers of tensors in a coherent way without representing the full mathematical detail (see Fig.~\ref{network}). For example, the wavefunction when decomposed through a series of reshapes and singular value decompositions (SVD) takes the form of a matrix product state (MPS) \cite{bakerCJP21}
\begin{align}\nonumber
|\psi\rangle&=\sum_{\substack{\sigma_1\sigma_2\sigma_3\\\sigma_4\sigma_5\sigma_6}}\sum_{\substack{a_1a_2a_3\\a_4a_5}}A^{\sigma_1}_{a_1}A^{\sigma_2}_{a_1a_2}A^{\sigma_3}_{a_2a_3}A^{\sigma_4}_{a_3a_4}A^{\sigma_5}_{a_4a_5}A^{\sigma_6}_{a_5}\\
&\quad\quad\quad\quad\quad\quad\quad\quad\times|\sigma_1\sigma_2\sigma_3\sigma_4\sigma_5\sigma_6\rangle
\label{mpsform}
\end{align}
which is far more intricate than the diagram notation but conveys no extra information. Any quantum operator (for example, the Hamiltonian) can be decomposed into a similar site-by-site form, called a matrix product operator (MPO).

@@

The library implements several classes of tensor network anstaz to represent wavefunctions of quantum problems. This includes the MPS \cite{verstraete2006matrix,baker2024bundled}, projected entangled pair states (PEPS) \cite{verstraete2004renormalization}, multiscale entanglement renormalization anstaz (MERA) \cite{vidal2007entanglement}, and classical tensor network algorithms such as the tensor renormalization group (TRG) \cite{levin2007tensor}. Each of these implementations is done with care towards ease of use, culminating in the DMRGenie implementation here.  All of these tensor networks are implemented in a fully flexible way.
